Investigating shape standardization using geometric morphometry of pottery from Iron Age northeastern Taiwan

### 2020-04-11 analysis/paper/paper.Rmd
editing figs

# Isodeformation lines
dev.off()

#knitr::include_graphics(here::here("analysis", "figures", "tps-pre-Euro-to-post-Euro.png"))
```

@@

```{r}
# PCA byplot, we save it to disk to use in the next chunk
png(here("analysis/figures/pca-chull.png"), res = 300, w = 1800, h = 1000)
plot_PCA(op.p, ~phase, chullfilled = TRUE)
dev.off()
```

```{r}
# composite composite image of thin plate spline and PCA output
iso_plot <- image_read(here("analysis/figures/tps-iso.png"))
iso_plot <- image_trim(iso_plot)
pca_plot <- image_read(here("analysis/figures/pca-chull.png"))

@@

```{r, first-two-PCs-byplot-and-tps-plot, echo=FALSE, fig.cap="caption", fig.align='Left: Comparison of average vessel shapes across phases using thin plate splines (TPS) with outline deformations required to pass from an extreme of one morphospace to another. Right: Pottery shape distribution by each occupation phase according to the first two PCs.'}
knitr::include_graphics(here("analysis/figures/iso_pca_composit_crop.png"))
```

@@

# join two dataset to get coordinate
pottery_count_join <-
  pottery_count %>%
  mutate(phase = as.factor(phase)) %>% # seems need this line first
  inner_join(AD_zone_tidy) %>% 
  filter(!is.na(phase)) %>% 
  mutate(phase = case_when(
    phase %in% "pre-e"~ "Before European Contact",
    phase %in% "post-e" ~ "European Presence",
    phase %in% "ch-con" ~ "Chinese Presence")) %>% 
  mutate(phase = factor(phase,
                        level = c("Before European Contact",
                                  "European Presence",
                                  "Chinese Presence")))
# plot spatial pattern for pottery
plot_spatial_distribution_pottery_by_phase <- 
  ggplot() +
  geom_sf(data = AD_zone_tidy, fill = NA) +
  geom_sf(data = pottery_count_join, 
          aes(geometry = geometry, 
              fill =n), 
          alpha = 0.9)+
  facet_wrap(~ phase) +
  labs(fill = "frequency") +
  scale_fill_viridis(direction = -1) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 6),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 6),
        legend.key.size =  unit(3, "mm"))

plot_spatial_distribution_pottery_by_phase

ggsave(plot = plot_spatial_distribution_pottery_by_phase,
       here("analysis", "figures",
            "plot-spatial-distribution-pottery-by-period.jpg"),
       width = 135, 
       height = 31,
       dpi = 300,
       units = "mm")
```

```{r kernel-plot, fig.cap= "Kernel density map visulizes the probability of the density of pottery across space. The map shows  major core area during the pre-European period, multiple core areas during the European period, and a single core during the Chinese period. Used the bandwidth based on Silverman (1986)'s rule of thumb"}
knitr::include_graphics(here("analysis", "figures", "plot-kde-maps.jpg"))
```

```{r Monte-Carlo-spatial-pattern, fig.cap= "Histograms of simulated average nearest-neighbour distances (ANN) values from 1000 simulations for three phases. X-axis values based on meteres represent ANN expected value. Each sample distribution presents the null hypothesis with the blue line indicating the observed ANN value"}
knitr::include_graphics(here("analysis", "figures", "plot-kde-ann-histograms.jpg"))
```

>>>>

# Isodeformation lines
invisible(dev.off())
```

@@

```{r}
# PCA byplot, we save it to disk to use in the next chunk
png(here("analysis/figures/pca-chull.png"), res = 300, w = 1800, h = 1000)
plot_PCA(op.p, ~phase, chullfilled = TRUE)
invisible(dev.off())
```

```{r}
library(magick)
# composite composite image of thin plate spline and PCA output
iso_plot <- image_read(here("analysis/figures/tps-iso.png"))
iso_plot <- image_trim(iso_plot)
pca_plot <- image_read(here("analysis/figures/pca-chull.png"))

@@

```{r, first-two-PCs-byplot-and-tps-plot, echo=FALSE, fig.cap='Left: Comparison of average vessel shapes across phases using thin plate splines (TPS) with outline deformations required to pass from an extreme of one morphospace to another. Right: Pottery shape distribution by each occupation phase according to the first two PCs.'}
knitr::include_graphics(here("analysis/figures/iso_pca_composit_crop.png"))
```

@@

# join two dataset to get coordinate
pottery_count_join <-
  pottery_count %>%
  mutate(phase = as.factor(phase)) %>% # seems need this line first
  inner_join(AD_zone_tidy) %>% 
  filter(!is.na(phase)) %>% 
  mutate(phase = case_when(
    phase %in% "pre-e"~ "Before European\nContact",
    phase %in% "post-e" ~ "European Presence",
    phase %in% "ch-con" ~ "Chinese Presence")) %>% 
  mutate(phase = factor(phase,
                        level = c("Before European\nContact",
                                  "European Presence",
                                  "Chinese Presence")))
# plot spatial pattern for pottery
plot_spatial_distribution_pottery_by_phase <- 
  ggplot() +
  geom_sf(data = AD_zone_tidy, fill = NA) +
  geom_sf(data = pottery_count_join, 
          aes(geometry = geometry, 
              fill =n), 
          alpha = 0.9)+
  facet_wrap(~ phase) +
  labs(fill = "frequency") +
  scale_fill_viridis(direction = -1) +
  theme_minimal() +
  theme(panel.spacing = unit(3.5, "lines"),
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        strip.text.x = element_text(size = 11),
       legend.text = element_text(size = 6),
       legend.title = element_text(size = 6, vjust = 0.75),
       legend.key.size =  unit(4, "mm"),
        legend.position = "top")

ggsave(plot = plot_spatial_distribution_pottery_by_phase,
       here("analysis", "figures",
            "plot-spatial-distribution-pottery-by-period.jpg"),
       width = 275, 
       height = 61,
       dpi = 300,
       units = "mm")
```

```{r kernel-plot, fig.cap= "A: The spatial distribution of the pottery selected for shape analysis. The quantity is indicated by the color scale. B: Kernel density map visulizes the probability of the density of pottery across space. The maps show a major core area during the pre-European period, multiple core areas during the European period, and a single core during the Chinese period. The bandwidth is based on Silverman (1986)"}

p1 <- ggdraw(clip = 'on') + 
  draw_image(here("analysis", "figures", "plot-spatial-distribution-pottery-by-period.jpg"), 
             scale = 1.6,
             x = -0.025)

# code for this is in the spatial_pattern.R script file
p2 <- ggdraw() + 
  draw_image(here("analysis", "figures", "plot-kde-maps.jpg"), 
             scale = 1)

plot_grid(p1, p2, ncol = 1, labels="AUTO")
ggsave(here("analysis/figures/kde-and-pottery-distribution-composite-plot.png"))
```

```{r Monte-Carlo-spatial-pattern, fig.cap= "Histograms of simulated average nearest-neighbour distances (ANN) values from 1000 simulations for three phases. X-axis values based on meteres represent ANN expected value. Each sample distribution presents the null hypothesis with the blue line indicating the observed ANN value", fig.width=3}

  image_resize(image_read( here("analysis", 
                                "figures", 
                                "plot-kde-ann-histograms.jpg")), 
               geometry_area(width = 1000))

```
